\textbf{SPECFEM3D Cartesian}\\ \textbf{User Manual}

### 2015-11-12 doc/USER_MANUAL/13_changing_the_model.tex
Buildbot ran the code beautifier.

\noindent
Each data record line provides the velocity model values in a format like:  
{\small
\begin{verbatim}
# data record format: purely isotropic model
x	y	z		vp		vs		rho
..
\end{verbatim}
}
\noindent
or
{\small
\begin{verbatim}
# data record format: anelastic isotropic model
x	y	z		vp		vs		rho	Qp	Qs
..
\end{verbatim}
}
\noindent
where  \texttt{x}, \texttt{y}, \texttt{z} are the grid point position, \texttt{vp}, \texttt{vs} and \texttt{rho} the P- and S-wave speeds and density, and \texttt{Qp} and \texttt{Qs} the quality factors for P- and S-wave speeds. The quality factors are optional and can be omitted for purely elastic models (the tomography routine will recognize both formats). Internally, the quality factors will be converted to bulk and shear attenuation values, 
$Q_{\kappa}$ and $Q_{\mu}$ respectively \citep{AnHa78}. 
For simulations with attenuation, please note that the Vp- and Vs-velocities of your model are given for a reference
frequency. To change this reference frequency, you change the value
of \texttt{ATTENUATION\_f0\_REFERENCE} in the main constants file
\texttt{constants.h} found in subdirectory \texttt{src/shared/}.

@@

To use your own external model, you can set in the \texttt{Par\_file}
the model parameter \texttt{MODEL = external}. Three-dimensional acoustic
and/or (an)elastic (attenuation) models may then be superimposed onto
the mesh based upon your subroutine in \texttt{model\_external\_values.f90}
located in directory \texttt{src/generate\_databases}. The call to
this routine would be as follows
\begin{verbatim}
call model_external_values(xmesh, ymesh, zmesh, &
            rho, vp, vs, qkappa_atten, qmu_atten, iflag_aniso, idomain_id)
\end{verbatim}
Input to this routine consists of:
\begin{description}
\item [{\texttt{xmesh,ymesh,zmesh}}] location of mesh point
\end{description}
Output to this routine consists of:
\begin{description}
\item [{\texttt{rho,vp,vs}}] isotropic model parameters for density $\rho$
($kg/m^{3}$), $v_{p}$ ($m/s$) and $v_{s}$ ($m/s$)
\item [{\texttt{qkappa\_atten,qmu\_atten}}] Bulk and shear wave quality factor: $0<Q_{\kappa,\mu}<9000$
\item [{\texttt{iflag\_aniso}}] anisotropic model flag, $0$ indicating
no anisotropy or $1$ using anisotropic model parameters as defined
in routine file \texttt{model\_aniso.f90}
\item [{\texttt{idomain\_id}}] domain identifier, $1$ for acoustic, $2$
for elastic, $3$ for poroelastic materials.
\end{description}
Note that the resolution and maximum value of anelastic models are
truncated. This speeds the construction of the standard linear solids
during the meshing stage. To change the resolution, currently at one
significant figure following the decimal, or the maximum value (9000),
consult \texttt{shared/constants.h}.

>>>>

\noindent
Each data record line provides the velocity model values in a format like:
{\small
\begin{verbatim}
# data record format: purely isotropic model
x y z   vp    vs    rho
..
\end{verbatim}
}
\noindent
or
{\small
\begin{verbatim}
# data record format: anelastic isotropic model
x y z   vp    vs    rho Qp  Qs
..
\end{verbatim}
}
\noindent
where  \texttt{x}, \texttt{y}, \texttt{z} are the grid point position, \texttt{vp}, \texttt{vs} and \texttt{rho} the P- and S-wave speeds and density, and \texttt{Qp} and \texttt{Qs} the quality factors for P- and S-wave speeds. The quality factors are optional and can be omitted for purely elastic models (the tomography routine will recognize both formats). Internally, the quality factors will be converted to bulk and shear attenuation values,
$Q_{\kappa}$ and $Q_{\mu}$ respectively \citep{AnHa78}.
Note that Qmu is always equal to Qs, but Qkappa is in general not equal to Qp. To convert one to the other see doc/note\_on\_Qkappa\_versus\_Qp.pdf and utils/attenuation/conversion\_from\_Qkappa\_Qmu\_to\_Qp\_Qs\_from\_Dahlen\_Tromp\_959\_960.f90.
For simulations with attenuation, please note that the Vp- and Vs-velocities of your model are given for a reference
frequency. To change this reference frequency, you change the value
of \texttt{ATTENUATION\_f0\_REFERENCE} in the main constants file
\texttt{constants.h} found in subdirectory \texttt{src/shared/}.

@@

To use your own external model, you can set in the \texttt{Par\_file}
the model parameter \texttt{MODEL = external}. Three-dimensional acoustic
and/or (an)elastic (attenuation) models may then be superimposed onto
the mesh based upon your subroutine in \texttt{model\_external\_values.f90}
located in directory \texttt{src/generate\_databases}. The call to
this routine would be as follows
\begin{verbatim}
call model_external_values(xmesh, ymesh, zmesh, &
            rho, vp, vs, qkappa_atten, qmu_atten, iflag_aniso, idomain_id)
\end{verbatim}
Input to this routine consists of:
\begin{description}
\item [{\texttt{xmesh,ymesh,zmesh}}] location of mesh point
\end{description}
Output to this routine consists of:
\begin{description}
\item [{\texttt{rho,vp,vs}}] isotropic model parameters for density $\rho$
($kg/m^{3}$), $v_{p}$ ($m/s$) and $v_{s}$ ($m/s$)
\item [{\texttt{qkappa\_atten,qmu\_atten}}] Bulk and shear wave quality factor: $0<Q_{\kappa,\mu}<9000$.
Note that Qmu is always equal to Qs, but Qkappa is in general not equal to Qp. To convert one to the other see doc/note\_on\_Qkappa\_versus\_Qp.pdf and utils/attenuation/conversion\_from\_Qkappa\_Qmu\_to\_Qp\_Qs\_from\_Dahlen\_Tromp\_959\_960.f90.
\item [{\texttt{iflag\_aniso}}] anisotropic model flag, $0$ indicating
no anisotropy or $1$ using anisotropic model parameters as defined
in routine file \texttt{model\_aniso.f90}
\item [{\texttt{idomain\_id}}] domain identifier, $1$ for acoustic, $2$
for elastic, $3$ for poroelastic materials.
\end{description}
Note that the resolution and maximum value of anelastic models are
truncated. This speeds the construction of the standard linear solids
during the meshing stage. To change the resolution, currently at one
significant figure following the decimal, or the maximum value (9000),
consult \texttt{shared/constants.h}.

### 2015-11-12 doc/USER_MANUAL/15_informations_for_developers.tex

\noindent We also have simple 3D demo source codes that implement the SEM in a single, small program, in directory\\
utils/small\_SEM\_solvers\_in\_Fortran\_and\_C\_without\_MPI\_to\_learn of the specfem3d package.
They are useful to learn how the spectral-element method works, and how to write or modify a code to implement it.
Also useful to test new ideas by modifying these simple codes to run some tests.\\

>>>>

\noindent We also have simple 3D demo source codes that implement the SEM in a single, small program, in directory\\
utils/small\_SEM\_solvers\_in\_Fortran\_and\_C\_without\_MPI\_to\_learn of the specfem3d package.
They are useful to learn how the spectral-element method works, and how to write or modify a code to implement it.
Also useful to test new ideas by modifying these simple codes to run some tests.
We also have a similar, even simpler, demo source code for the 2D case in directory\\
utils/small\_SEM\_solver\_in\_Fortran\_without\_MPI\_to\_learn of the specfem2d package.
